<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{R/simcross User Guide}
-->

# aRxiv tutorial

[arXiv](http://arxiv.org) is an archive of manuscripts for computer
science, mathematics, physics, quantiative biology, quantitative
finance, and statistics. The
[aRxiv package](https://github.com/ropensci/aRxiv) provides an
[R](http://www.r-project.org) interface to the
[arXiv API](http://arxiv.org/help/api/index).

Note that the arXiv API _does not_ require an API key.

```{r change_aRxiv_delay_option, include=FALSE}
options(aRxiv_delay=0.5)
```

## Installation

The [aRxiv package](https://github.com/ropensci/aRxiv) is not
currently available on [CRAN](http://cran.r-project.org). To install the
package from [GitHub](http://github.com), use
`devtools::install_github()`, as follows:

```coffee
install.packages("devtools")
library(devtools)
install_github("ropensci/aRxiv")
```

## Basic use

The aRxiv package currently includes two functions: `arxiv_search` for
general search, and `arxiv_count` for a simple count of the number
of manuscripts matching a given search query.

The two basic arguments are `query`, a
character string representing the search, and `id_list`, a list of
[arXiv manuscript identifiers](http://arxiv.org/help/arxiv_identifier). 

- If only `query` is provided, manuscripts matching that query are
  returned
- If only `id_list` is provided, manuscripts in the list are
  returned
- If both are provided, manuscripts in `id_list` that match `query`
  will be returned.

`query` may be a single character string or a vector of character
strings. If it is a vector, the elements are pasted together with
`AND`.

`id_list` may be a vector of strings or a single
comma-separated character string.

Here's a simple example, search for manuscripts with the phrase
`"Peter Hall"` in one of the author fields.

```{r first_search}
library(aRxiv)
arxiv_count('au:"Peter Hall"')
z <- arxiv_search('au:"Peter Hall"', limit=50)
nrow(z)
```


## Counting papers

## Forming queries

## Subject classification


## Technical details

### Metadata limitations

### Limit time between search requests

### Limit number of items returned

### Requests in batches

## License and bugs

- License:
  [MIT](https://github.com/ropensci/aRxiv/blob/master/LICENSE)
- Report bugs or suggestions improvements by [submitting an issue](https://github.com/ropensci/aRxiv/issues) to
  [our GitHub repository for aRxiv](https://github.com/ropensci/aRxiv).



<!-- the following to make it look nicer -->
<link href="http://kbroman.org/qtlcharts/assets/vignettes/vignette.css" rel="stylesheet"></link>
